02_DESeq_characterization

### 2019-08-26 analysis/03_DESeq2_characterization.Rmd
updated characterization analysis

```{r dds, cache = T, eval = F}
dds.female_hypothalamus <- subsetDESeq(c.colData,  c.countData, "female_hypothalamus")
dds.female_pituitary <- subsetDESeq(c.colData,  c.countData, "female_pituitary" )
dds.female_gonad <- subsetDESeq(c.colData,  c.countData, "female_gonad" )
dds.male_hypothalamus <- subsetDESeq(c.colData,  c.countData, "male_hypothalamus" )
dds.male_pituitary <- subsetDESeq(c.colData,  c.countData, "male_pituitary"  )
dds.male_gondad <- subsetDESeq(c.colData,  c.countData, "male_gonad")
```

@@

```{r totalDEGs, eval = F}
#create list of groups for deseq contrasts
group1 <- c("control",  "bldg", "lay", "inc.d3", "inc.d9", "inc.d17", "hatch", "n5", "n9")
group2 <- group1

a <- plottotalDEGs(dds.female_hypothalamus, "female hypothalamus")
b <- plottotalDEGs(dds.female_pituitary, "female pituitary")
c <- plottotalDEGs(dds.female_gonad, "female gonad")
d <- plottotalDEGs(dds.male_hypothalamus, "male hypothalamus")
e <- plottotalDEGs(dds.male_pituitary, "male pituitary")
f <- plottotalDEGs(dds.male_gondad, "male gonad")

@@

@@

```{r pca, eval = F}
plotPCAs(dds.female_hypothalamus, "female hypothalamus")
plotPCAs(dds.female_pituitary, "female pituitary")      
plotPCAs(dds.female_gonad, "female gonad")
plotPCAs(dds.male_hypothalamus, "male hypothalamus")
plotPCAs(dds.male_pituitary, "male pituitary")
plotPCAs(dds.male_gondad, "male gonad")
```

>>>>

```{r dds, cache = T, eval = T}
dds.female_hypothalamus <- subsetDESeq(c.colData,  c.countData, "female_hypothalamus")
dds.female_pituitary <- subsetDESeq(c.colData,  c.countData, "female_pituitary" )
dds.female_gonad <- subsetDESeq(c.colData,  c.countData, "female_gonad" )
dds.male_hypothalamus <- subsetDESeq(c.colData,  c.countData, "male_hypothalamus" )
dds.male_pituitary <- subsetDESeq(c.colData,  c.countData, "male_pituitary"  )
dds.male_gondad <- subsetDESeq(c.colData,  c.countData, "male_gonad")
```

@@

```{r returntotalDEGs, eval = T}
DEGs.female_hypothalamus <- returntotalDEGs(dds.female_hypothalamus)
DEGs.female_pituitary <- returntotalDEGs(dds.female_pituitary)
DEGs.female_gonad <- returntotalDEGs(dds.female_gonad)
DEGs.male_hypothalamus <- returntotalDEGs(dds.male_hypothalamus)
DEGs.male_pituitary <- returntotalDEGs(dds.male_pituitary)
DEGs.male_gondad <- returntotalDEGs(dds.male_gondad)
```

```{r totalDEGs, eval = T}
#create list of groups for deseq contrasts
group1 <- c("control",  "bldg", "lay", "inc.d3", "inc.d9", "inc.d17", "hatch", "n5", "n9")
group2 <- group1

a <- plottotalDEGs(DEGs.female_hypothalamus, "female hypothalamus")
b <- plottotalDEGs(DEGs.female_pituitary, "female pituitary")
c <- plottotalDEGs(DEGs.female_gonad, "female gonad")
d <- plottotalDEGs(DEGs.male_hypothalamus, "male hypothalamus")
e <- plottotalDEGs(DEGs.male_pituitary, "male pituitary")
f <- plottotalDEGs(DEGs.male_gondad, "male gonad")

@@

## Plot only a subset of DEGs that relate to hypotheses as bar and line graphs

```{r barplots}
# bars with lines
g <- plotserialDEGs(DEGs.female_hypothalamus, "Female hypothalamus", "#00A08A")
h <- plotserialDEGs(DEGs.female_pituitary, "Female pituitary", "#00A08A")
i <- plotserialDEGs(DEGs.female_gonad, "Female gonad", "#00A08A")
j <- plotserialDEGs(DEGs.male_hypothalamus, "Male hypothalamus", "#F98400")
k <- plotserialDEGs(DEGs.male_pituitary, "Male pituitary", "#F98400")
l <- plotserialDEGs(DEGs.male_gondad, "Male gonad", "#F98400")

mybarplots <- plot_grid(g, h , i , j ,k , l, nrow = 2) 
mybarplots

pdf("../figures/characterization/barplots.pdf", width = 10, height = 6)
plot(mybarplots)
dev.off()
```

```{r linegraphs}

# lines only, different colors for different sex tissue groups
FH <- subsetDEGs(DEGs.female_hypothalamus, "Female hypothalamus")
FP <- subsetDEGs(DEGs.female_pituitary, "Female pituitary")
FG <- subsetDEGs(DEGs.female_gonad, "Female gonad")
MH <- subsetDEGs(DEGs.male_hypothalamus, "Male hypothalamus")
MP <- subsetDEGs(DEGs.male_pituitary, "Male pituitary")
MG <- subsetDEGs(DEGs.male_gondad, "Male gonad")

specificDEGs <- rbind(FH, FP, FG, MH, MP, MG)

specificDEGs$sextissue <- factor(specificDEGs$sextissue, levels = c("Female hypothalamus", "Female pituitary", "Female gonad",
                                                                    "Male hypothalamus", "Male pituitary", "Male gonad"))
mylinegraph <- ggplot(specificDEGs, aes(x=comparison, y=V3, group = sextissue)) +
    geom_line(aes(color=sextissue)) +
    geom_point(aes(color=sextissue)) + 
    theme_minimal(base_size = 10) +
    #theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(x = "Parental stage transitions", y= "No. of differentially expressed genes ",
         title = "Gene expression changes during parental care transitions") + 
    scale_x_discrete(labels=c("Control to\nNest bldg", "Nest bldg. to\nLay", "Lay to\nInc. day 3",
                              "Inc. day 3 to\nInc. day 9", "Inc. day 9 to\nInc. day 17",
                              "Inc. day 17 to\nHatch", "Hatch to\nNestling day 5", "Nestling day 5 to\n Nestling day 9")) +
   theme(legend.title = element_blank(),
         legend.position = c(0.8, 0.8))
mylinegraph

pdf("../figures/characterization/linegraphs.pdf", width = 8, height = 4)
plot(mylinegraph)
dev.off()
```

@@

```{r pca, eval = T}
FhypPCA <- returnPCAs(dds.female_hypothalamus)
FpitPCA <- returnPCAs(dds.female_pituitary)      
FgonPCA <- returnPCAs(dds.female_gonad)
MhypPCA <- returnPCAs(dds.male_hypothalamus)
MpitPCA <- returnPCAs(dds.male_pituitary)
MgonPCA <- returnPCAs(dds.male_gondad)

plotPC12(FhypPCA, "female hypothalamus")
plotPC12(FpitPCA, "female pituitary")
plotPC12(FgonPCA, "female gonad")
plotPC12(MhypPCA, "male hypothalamus")
plotPC12(MpitPCA, "male pituitary")
plotPC12(MgonPCA, "male gonad")
```
